You have more money in your wallet. Does that mean you will spend extra on food?

### 2024-04-08 paper/paper.qmd
finished introduction and abstract

---
title: "You have more money in your wallet. Does that mean you will spend extra on food?"
subtitle: "I analyze the relationship between disposable income and food expenditure for American consumers"
author: Ahnaf Alam
thanks: "Code and data are available at https://github.com/AhnafAlam1/income_and_expenditure"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

You can and should cross-reference sections and sub-sections. We use @citeR and @rohan.

The remainder of this paper is structured as follows. @sec-data....

@@

# Model

@@

# Results

@@

# Discussion

>>>>

---
title: "You have more money in your wallet. Does that mean you will spend extra on food?"
subtitle: "I analyze the relationship between disposable income and food expenditure for American consumers"
author: Ahnaf Alam
thanks: "Code and data are available at https://github.com/AhnafAlam1/income_and_expenditure"
date: today
date-format: long
classoption: abstract
abstract: "We estimated how food spending changes when disposable income increases. Using both a simple linear regression model and a multiple regression model, we found an economically insignificant increase of 20 million USD for every billion dollar increase in disposable income. At a policy level, our conclusion suggests that focusing solely on disposable income will not significantly affect food spending patterns. Policies aimed at stimulating consumer spending must consider other avenues."
format: pdf
number-sections: true
bibliography: references.bib
toc: true
---

@@

Wages have been going up in the US for the past thirty years, which usually means people have more money to spend after taxes. This extra money, called disposable income, is what's left over after taxes are taken out of a person's income. It's an important thing to look at because it shows how much money people have to spend. Research shows that when people have more disposable income, they tend to spend more money [@friend1946relationship]. We're interested in how strong this connection is when it comes to one specific thing: food.

Food is something everyone needs to live. People might change what kinds of food they like over time, but everyone still needs to eat. In this project, we're looking at how closely linked disposable income is to how much people spend on food. This matters because if there's a strong connection, it suggests that when people make more money, they tend to spend more on basics like food. This could help predict how much people will spend on luxury items in the future. For instance, if we find a strong connection, it might mean that when people earn more, they spend a bigger chunk of their money on essentials like food, leaving less for luxury items.

Numerous studies, notably summarized by Ernst Engel's law, suggest that as income and family size increase, spending on food goes up, but the proportion of income spent on food decreases [@lewbel2008engel]. This means wealthier families tend to spend more on food, but it's a smaller part of their overall income. Taking inspiration from Engel's law, we want to see how food spending changes when we look at other spending categories instead of just income differences across families.

Additionally, most past studies used income as the predictor variable, but we're using disposable income because we think it gives a better idea of overall wealth. The estimand of this paper is the measure of how much food spending changes (in billions of USD) when disposable income increases by a billion USD.

In @sec-data, we discuss the data we used for this project, including information of the software and packages used. This is followed by a explanation of the models that we used, which can be found in @sec-model. In @sec-res, we present the results of the models and @sec-dis, we interpret these results, and consider how savings, consumer preference and elasticies play into this discussion on food expenditure and income. We finish the project with weaknesses and potential areas of improvement in the future.

@@

# Model {#sec-model}

@@

# Results {#sec-res}

@@

# Discussion {#sec-dis}
